\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-19 redaction/dhp_july2017.tex
changement

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}
\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille1.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
This paper provides a quantitative assessment of the size and the importance of the additive component
in international transport costs. To do so, we exploit information contained in the US imports flows database over 1974-2013, distinguishing by transport mode, air and vessel. Two main new findings emerge. First, the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. Second, and in contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer.
\end{abstract}

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel, in line with \cite{hummels2007}. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components. In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. In quantitative terms further, we find that the additive component of transport costs is sizeable, with mean values over the period of 1.8\% and 2.9\% of the export price in air and vessel transport respectively. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

To prove this point drives us to provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Rigourously speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts. To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over the more recent years through (2005-2013). \smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs. We obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the size of international transport costs. Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

\textbf{VV: Affirmer nettement la difference avec Hummels !! Pour Vincent, notre vrai resultat est la (il y a une baisse des couts de transport, les effets de compo n’expliquent pas tout), et c’est cela que nous devrions vendre, davantage que l’histoire multiplicatif/additif, une question de geek en trade qui limitera forcément le type de revue auquel on pourra soumettre. EST-CE BON AINSI?}

\textbf{AI CHANGE LES SECTIONS}

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Considered in logs, the above equation becomes:

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e. on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively. This would not solve the problem though, for three main complementary reasons. First, at the 5-digit level the number of fixed effects to include in the estimation would be \textbf{very high - UN PEU MOU, dire combien?}, making the estimation computationally extremely burdensome, even in OLS. Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects. Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$. Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, rigourously speaking, the precise equation that we estimate is the following:
$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode). For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5 percent of the upper and lower tails of the distribution in the regression variables. These cut-offs are aimed at eliminating reporting or coding errors. We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$). Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.   \medskip

\textbf{VV: Les explications accompagnant l’equation estimee 6 ne sont pas claires : si $\tau_i$, $tau_{s(k)}$, $t_i$ et $t_{s(k)}$ sont des effets fixes, alors $tau_i = t_i$ et $tau_{s(k)} = t_{s(k)}$ !!! Clarifier. Cf FOOTNOTE AJOUTEE? EST-CE OK AINSI?}

\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu il y a quand meme de la variation between sectors au niveau 5-digits. Or, nous reagrigeons au niveau 3 ou 4d a cause des NLLS. Pourquoi ne pas estimer l'equation initiale (juste avant la (5)) en niveau avec des OLS ? A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante. AJOUTE UNE FOOTNOTE ET MODIFIE FACON DE JUSTIFIER FORME ERREUR MAIN TEXT. EST-CE OK AINSI?}

@@

One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price). However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis). Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control. In this respect, we are confident that our estimation strategy is immune from endogeneity problems.

\textbf{pas encore completement clair. Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde? Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement? Si c'est ca, c'est vrai seulement si les couts sont advalorem... Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus? Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip

>>>>

\title{\textsc{International Transport costs:\\New Findings from modeling additive costs}\thanks{We are grateful to participants at ETSG 17\textsuperscript{th} meetings in Helsinki, DEGIT XXII in Paris School of Economics, and participants at seminars of the University of California-Irvine, the Universities of Paris-Dauphine Paris-I and Nancy for very useful discussions. Finally, any remaining errors are ours. This paper features an online appendix containing additional results and available on the authors' websites.}}

\author{Guillaume \textsc{Daudin}\thanks{%
Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE \newline
Universit\'{e} Paris-Dauphine, PSL Research University, IRD, LEDa, UMR 225, DIAL, 75016 PARIS, FRANCE \newline
Sciences Po, Observatoire Fran\c{c}ais des Conjonctures \'{E}conomiques (OFCE), 75014 PARIS, FRANCE \newline
; email: \url{guillaume.daudin@dauphine.fr}}  \qquad J\'{e}r\^{o}me \textsc{H\'{e}ricourt} \thanks{Universit\'{e} de Lille - LEM-CNRS UMR 9221, France \& CEPII, France; email: \url{jerome.hericourt@univ-lille.fr}}\qquad Lise \textsc{Patureau}\thanks{Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE;  email: \url{lise.patureau@dauphine.fr} } }

@@

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel, in line with \cite{hummels2007}. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components. In contrast to \cite{hummels2007}, we find that composition effects play a minor role in accounting for the reduction of transport costs relative to the value of US imports. Transport cost ``per se'' did decline. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

To prove this point drives us to provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded.\footnote{As documented by \cite{Irrazabal_2015}, pricing structure in shipping, additive tariffs, distribution costs... often exhibit (at least partly) an additive structure.} Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts. To do so, we exploit information contained in the US Imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over the more recent years through (2005-2013). \smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data only allows recovering the ratio of additive to ad-valorem costs as a fraction of export price; in contrast, our empirical framework provides a distinct measure of each additive and ad-valorem cost component. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. Based on the estimated share of transport costs in total trade costs provided by \citet{anderson_wincoop_jel} (around 15\%), straightforward calculation yields a ratio of additive to multiplicative transport (and not trade) costs of 2.1\% (14\% $\times$ 15\%) in \citet{Irrazabal_2015}. This falls in the range of our own estimates, and makes us confident in our quantitative assessment of the size of international transport costs. \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows a normal law centered on 0. Considered in logs, the above equation becomes:

@@

The non-linearity of Equation (\ref{eq:equation0}) implies that it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next. See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Confronted to this arbitrage, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at the 5-digit classification level ($k$). In other words, this amounts making the additional assumption, that all products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price. As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sector ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e. on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively. This would not solve the problem though, for three main complementary reasons. First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS. Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects. Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$. Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode). For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5 percent of the upper and lower tails of the distribution in the regression variables. These cut-offs are aimed at eliminating reporting or coding errors. We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation mode reported (air or vessel), on a sectoral-origin country basis ($i,s$). Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.   \medskip

@@

One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price). This is not an issue here, as we are not interested in causal inference. Rather, our aim is to provide an accounting breakdown of transport costs between the additive and the multiplicative components.

%However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis). Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control. In this respect, we are confident that our estimation strategy is immune from endogeneity problems.

%\textbf{pas encore completement clair. Veut-on dire que ce qui est pbtique, c'est que TC reagisse a ptilde? Je crois que ce qu'on veut dire, c'est que le fait que ptilde change en fonction des TC ne change pas les TC, le cif price s'ajuste proportionnellement? Si c'est ca, c'est vrai seulement si les couts sont advalorem... Pour dire que les firmes peuvent difficilement influencer le prix du transport au niveau 3d, on pourrait donner des stats des la dessus? Par produit 3d/pays, par secteur 3d tous pays, etc...?}\smallskip
